Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/evaluation.tex
Foo

With an understanding of which guarantees are and are not
HAT-compliant, in this section, we analyze the implications of these
results for existing systems and applications and briefly study HAT
systems on public cloud infrastructure. Specifically:

\begin{myenumerate}
\item We revisit traditional database concurrency control with a focus
  on coordination costs and on high availability.
\item We examine the properties required by a realistic OLTP
  application patterned on the TPC-C benchmark.
\item We perform a brief experimental evaluation of HAT versus non-HAT
  properties on public cloud infrastructure.
\end{myenumerate}

@@

Many existing database transaction and concurrency control algorithms
are not designed for high availability. These algorithms often presume
a single-server deployment or the requirement for serializability. As
a consequence, traditional transaction processing systems are not
well-optimized for a HAT context. In this section, we briefly discuss
design decisions and algorithmic details that preclude high
availability.

\vspace{.5em}\noindent\textbf{Serializability} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} may require at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ requires $T$ \texttt{lock} operations
and at least one \texttt{unlock} operation.  In a distributed
environment, each of these operations requires coordination, either
with other database replicas or with a lock service. If this
coordination mechanism is unavailable, transactions cannot safely
commit. Similarly, optimistic concurrency control requires
coordinating via a validation step, while deterministic transaction
scheduling~\cite{deterministic-scheduling} requires a
scheduler. Serializability under multi-version concurrency control
also requires checking for update conflicts. The reliance on a
globally agreed total order necessitates a minimum of one round-trip
to a designated master or coordination service for each of these
classic algorithms.  The cost of the (minimum one) round trip will be
determined by the deployment environment, as we saw in
Section~\ref{sec:motivation}; we will demonstrate this cost on public
cloud infrastructure in the next section.

@@

To better understand the impact of HAT versus non-HAT algorithms and
validate the latency-based analysis in Section~\ref{sec:latency}, we
deployed a basic distributed database prototype on public cloud
infrastructure. We compared the behavior of eventual consistency,
per-key mastering (wherein all operations required a RTT to a
designated master server), Read Committed, and TA with read your
writes guarantees. We deployed an experimental Java-based distributed
datastore backed by LevelDB both within and across EC2 regions and ran
the YCSB benchmark~\cite{ycsb} to determine the impact of latency. We
partitioned the database into two fully replicated clusters of three
\texttt{m1.xlarge} instances. For HAT, we stuck YCSB clients to their
local cluster and, to provide transactional functionality, grouped
every six YCSB operations from the default workload to form a
transaction. We report the median of five trials per experiment.

>>>>

With an understanding of which guarantees are HAT-compliant, in this
section, we analyze the implications of these results for existing
systems and briefly study HAT systems on public cloud
infrastructure. Specifically:

\begin{myenumerate}
\item We revisit traditional database concurrency control with a focus
  on coordination costs and on high availability.
\item We examine the properties required by a realistic OLTP
  application based on the TPC-C benchmark.
\item We perform a brief experimental evaluation of HAT versus non-HAT
  properties on public cloud infrastructure.
\end{myenumerate}

@@

Many existing database transaction and concurrency control algorithms
are not designed for high availability. These algorithms often presume
a single-server deployment or a requirement for serializability. As
a consequence, traditional transaction processing systems are not
well-optimized for a HAT context. In this section, we briefly discuss
design decisions and algorithmic details that preclude high
availability.

\vspace{.5em}\noindent\textbf{Serializability} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} may require at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ may require $T$ \texttt{lock}
operations and will require at least one \texttt{lock} and
\texttt{unlock} operation.  In a distributed environment, each of
these operations requires coordination, either with other database
replicas or with a lock service. If this coordination mechanism is
unavailable, transactions cannot safely commit. Similarly, optimistic
concurrency control requires coordinating via a validation step, while
deterministic transaction scheduling~\cite{deterministic-scheduling}
requires a scheduler. Serializability under multi-version concurrency
control also requires checking for update conflicts. The reliance on a
globally agreed total order necessitates a minimum of one round-trip
to a designated master or coordination service for each of these
classic algorithms.  The cost of the (minimum one) round trip will be
determined by the deployment environment, as we saw in
Section~\ref{sec:motivation}; we will demonstrate this cost on public
cloud infrastructure in the next section.

@@

To better understand the performance implications of HAT guarantees in
a real-world environment and to validate the latency-based analysis in
Section~\ref{sec:latency}, we deployed a distributed database
prototype on public cloud infrastructure. We compared the behavior of
eventual consistency, per-item mastering (wherein all operations to a
given item required contacting a designated master server), Read
Committed, and TA with read your writes guarantees. We deployed an
experimental Java-based distributed datastore backed by LevelDB both
within and across EC2 regions and ran the YCSB benchmark~\cite{ycsb}
to determine the impact of latency. We partitioned the database into
two fully replicated clusters of three \texttt{m1.xlarge} instances
each. For HAT, we stuck YCSB clients to their local cluster and, to
provide transactional functionality, grouped every six YCSB operations
from the default workload to form a transaction. We report the median
of five trials per experiment.

### 2013-04-01 writeup/intro.tex
Foo

Indeed, serializable transactions, the gold standard of traditional
ACID databases, are indeed not achievable given a requirement for high
availability in the presence of
partitions~\cite{davidson-survey}. However, database systems have a
long tradition of providing weaker isolation and consistency
guarantees~\cite{adya, ansicritique, gray-virtues, gray-isolation,
  kemme-thesis}. Due to their concurrency and overall performance
benefits, weak isolation models are widely employed, overwhelmingly as
the default setting in today's traditional ``ACID'' and up-and-coming
``NewSQL'' databases and often as the maximum guarantee offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics of current highly available systems. This raises a natural
question: which of these alternative semantics can be provided while
remaining highly available?

>>>>

Indeed, serializable transactions---the gold standard of traditional
ACID databases---are not achievable given a requirement for high
availability in the presence of
partitions~\cite{davidson-survey}. However, database systems have a
long tradition of providing weaker isolation and consistency
guarantees~\cite{adya, ansicritique, gray-virtues, gray-isolation,
  kemme-thesis}. Due to their concurrency and overall performance
benefits, weak isolation models are widely employed, overwhelmingly as
the default setting in today's traditional ``ACID'' and up-and-coming
``NewSQL'' databases and often as the maximum guarantee offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics of current highly available systems. This raises a natural
question: which of these alternative semantics can be provided while
remaining highly available?

### 2013-04-01 writeup/model.tex
Foo

Traditionally, a system provides {\em  availability} if every user that
can contact a correct server eventually receives a response from that server,
even in the presence of arbitrary, indefinitely long network
partitions between servers~\cite{gilbert-cap}. As in a standard
distributed database, designated servers might handle operations on
different data items;\footnote{There is a further distinction between
  a \textit{fully replicated} system, in which all servers are
  replicas for all data items, and a \textit{partially replicated}
  system, in which at least one server acts as a replica for only a
  subset of all data items. For generality, and, for applicability to
  many modern ``sharded'' or ``partitioned'' data storage
  systems~\cite{dynamo, pnuts, bigtable, spanner, hstore}, we consider partially
  replicated systems.} a server that can handle an operation for a
given data item is called a \textit{replica} for that item. While this
standard model is useful, there are several other cases we should
consider.

>>>>

Traditionally, a system provides {\em availability} if every user that
can contact a correct server eventually receives a response from that
server, even in the presence of arbitrary, indefinitely long network
partitions between servers~\cite{gilbert-cap}.\footnote{Under this
  definition, a system that requires a majority of servers to be
  online is not availabile.} As in a standard distributed database,
designated servers might handle operations on different data
items;\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{dynamo,
    pnuts, bigtable, spanner, hstore}, we consider partially
  replicated systems.} a server that can handle an operation for a
given data item is called a \textit{replica} for that item. While this
standard model is useful, there are several other cases we should
consider.
